Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization

### 2020-11-18 latex/slides/slides.tex
finished pass on experimental slides

\begin{frame}{Experimental Data Sets}
    \begin{itemize}
        \item DC-HYDICE
        \begin{itemize}
            \item Hyperspectral Digital Imagery Collection Experiment (HYDICE) of the National Mall in Washington, DC
        \end{itemize}
        \vfill
        \item SIIM-ISIC
        \begin{itemize}
            \item Society for Imaging Informatics in Medicine - International Skin Imaging Colloboration image classification of melanoma images
        \end{itemize}
        \vfill
        \item Synthetic Image classification
        \begin{itemize}
            \item smaller image classification dataset which has the same aspect ratio as SIIM-ISIC but fits in one node's memory
        \end{itemize}
    \end{itemize}
    \vfill
    \begin{center}
    \scriptsize
    	All experiments performed on Summit at Oak Ridge Leadership Computing Facility
    \end{center}
\end{frame}

\begin{frame}{R2NMF Strong Scaling}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_rank2_speedup.pdf}
            \caption{Synthetic  Data}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/realworld_rank2_speedup.pdf}
            \caption{\image{} Data}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{R2NMF Scaling Time Breakdowns}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_rank2_strongscaling.pdf}
            \caption{Synthetic  Data}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/realworld_rank2_strongscaling.pdf}
            \caption{\image{} Data}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{HierNMF Strong Scaling}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_hierarchical_speedup.pdf}
            \caption{Synthetic  Data}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/realworld_hierarchical_speedup.pdf}
            \caption{\image{} Data}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{HierNMF Scaling Time Breakdowns}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_hier_strongscaling.pdf}
            \caption{Synthetic  Data}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/realworld_hier_strongscaling.pdf}
            \caption{\image{} Data}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Tree Level Times on Synthetic Data}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_sequential_level_breakdown.pdf}
            \caption{1 Compute Node}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_parallel_level_breakdown.pdf}
            \caption{40 Compute Nodes}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Rank Scaling for Hierarchical and Flat NMF}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/dc_rankscaling.pdf}
            \caption{\hyper{}  Data}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_rankscaling.pdf}
            \caption{Synthetic Data}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

>>>>

\begin{frame}{Experimental Data Sets}
    \begin{itemize}
        \item DC-HYDICE (191 $\times$ 392{,}960, 600MB)
        \begin{itemize}
            \item Hyperspectral Digital Imagery Collection Experiment (HYDICE) of the National Mall in Washington, DC
        \end{itemize}
        \vfill
        \item SIIM-ISIC (3{,}145{,}728 $\times$ 33{,}126, 800GB)
        \begin{itemize}
            \item Society for Imaging Informatics in Medicine - International Skin Imaging Colloboration image classification of melanoma images
        \end{itemize}
        \vfill
        \item Synthetic (1{,}048{,}576 $\times$ 11{,}042, 86GB)
        \begin{itemize}
            \item smaller synthetic image classification dataset which has the same aspect ratio as SIIM-ISIC but fits in one node's memory
        \end{itemize}
    \end{itemize}
    \vfill
    \begin{center}
    \scriptsize
    	All experiments performed on Summit at Oak Ridge Leadership Computing Facility
    \end{center}
\end{frame}

\newcommand{\figscal}{.75\textwidth}

\begin{frame}{R2-NMF Strong Scaling and Time Breakdown}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/synthetic_rank2_speedup.pdf}
            \caption{Synthetic Scaling}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/realworld_rank2_speedup.pdf}
            \caption{\image{} Scaling}
            \end{figure}
        \end{column}
    \end{columns}
\vspace{-.5cm}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/synthetic_rank2_strongscaling.pdf}
            \caption{Synthetic Breakdown}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/realworld_rank2_strongscaling.pdf}
            \caption{\image{} Breakdown}
            \end{figure}
        \end{column}
    \end{columns}

\begin{itemize}
\small
    \item Nearly perfect strong scaling because time dominated by local matrix multiplication
\end{itemize}

\begin{frame}{Hierarchical Clustering Scaling and Breakdown}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/synthetic_hierarchical_speedup.pdf}
            \caption{Synthetic  Data}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/realworld_hierarchical_speedup.pdf}
            \caption{\image{} Data}
            \end{figure}
        \end{column}
    \end{columns}
\vspace{-.5cm}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/synthetic_hier_strongscaling.pdf}
            \caption{Synthetic  Data}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\figscal]{../plots/realworld_hier_strongscaling.pdf}
            \caption{\image{} Data}
            \end{figure}
        \end{column}
    \end{columns}

\begin{itemize}
\small
    \item Scaling hindered by NNLS computations and communication of small subproblems at leaves of hierarchy
\end{itemize}

\begin{frame}{Tree Level Times on Synthetic Data}
    \centering
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_sequential_level_breakdown.pdf}
            \caption{1 Compute Node}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \includegraphics[width=\textwidth]{../plots/synthetic_parallel_level_breakdown.pdf}
            \caption{40 Compute Nodes}
            \end{figure}
        \end{column}
    \end{columns}

\begin{itemize}
    \item Evidence that scaling is hindered by leaves of tree
    \begin{itemize}
    	\item MatMul costs are constant over levels of balanced tree 
	\item NNLS and communication costs grow over levels
    \end{itemize}
\end{itemize}

\begin{frame}{Conclusions and Future Directions}

\begin{itemize}
	\item Scalability is possible when MatMul dominates, particularly when input data is tall and skinny and tree is balanced
	\vfill
	\item 2D distribution of matrix $\M{A}$ could improve scalability but requires redistribution of $\M{A}$ for subtrees
	\vfill
	\item Parallelization across tree could improve scalability but requires careful load balancing and decentralizing split orders
	\vfill
	\item Hierarchical clustering can also be used to initialize flat NMF, particularly when $k$ is large, for Divide-and-Conquer NMF
\end{itemize}
